### 2015-02-23 fundamentals/impl.tex
move more into caption

In large simulations, many material objects may change frequently.  Material
decay can also contribute significantly to such changes.  In order to help
avoid unnecessary runtime performance and database size impacts, compositions
in \Cyclus have some special features.  In particular, compositions are
immutable once created.  This allows multiple material objects to hold
references to the same composition safely.  Additionally, new compositions
resulting from decay are cached and used to avoid redundant decay
calculations.  Composition immutability in concert with decay history caching
help eliminate many redundant calculations in addition to reducing the
total number of composition entries recorded in the database.

Figure \ref{fig:compositions} shows a set of four compositions A, B, C, and a
yet-uncomputed decayed composition D.  B comes from decaying A 1 time step.  C
comes from decaying B 2 time steps, etc.  Black represents the decay history
cache for this particular composition chain. Blue represents decay
calculations that are satisfied by cache lookups.   If A needs to be decayed 1
time step (A +1), a lookup is performed on the decay line and B is found and
returned - no decay calculation is performed and no new compositions need to
be created or recorded.  Decaying B 3 time steps requires a decay calculation
to compute D, but subsequent requests such as decaying A 4 time steps will not
require any recalculation.

\begin{figure}[htbp!]
\begin{center}
\includegraphics{./images/compositions}
\end{center}
\caption{Composition decay history chain helps avoid redundant decay calculations and reduce duplication in the database.}
\label{fig:composition}
\end{figure}

>>>>

In large simulations, many material objects may change frequently.  Material
decay can also contribute significantly to such changes.  In order to help
avoid unnecessary runtime performance and database size impacts, compositions
in \Cyclus have some special features.  In particular, compositions are
immutable once created.  This allows multiple material objects to hold
references to the same composition safely.  Additionally, new compositions
resulting from decay are cached and used to avoid redundant decay
calculations.  Figure \ref{fig:compositions} illustrates how this decay
history cache works. Composition immutability in concert with decay history
caching help eliminate many redundant calculations in addition to reducing the
total number of composition entries recorded in the database.

\begin{figure}[htbp!]
\begin{center}
\includegraphics{./images/compositions}
\end{center}
\caption{
A simple decay line cache holding compositions A, B, C, and a
yet-uncomputed composition D.  B comes from decaying A 1 time step.  C comes
from decaying B 2 time steps, etc.  Black represents the cache for this
particular composition chain. Blue indicates decay operations that can be
satisfied by cache lookups.   If A needs to be decayed 1 time step (A +1), a
quick lookup returns the previously computed B.  Decaying B 3 time steps
requires a decay calculation to compute a new composition D, but subsequent
requests such as decaying A 4 time steps will not require any recalculation.
}
\label{fig:compositions}
\end{figure}
